---
title: "Bellabeat_Casestudy"
author: "Aparna"
date: '2022-05-26'
output: html_document
---

##Installing the packages.
Install the packages as and when needed.

```{r}
# Install necessary packages.

install.packages("tidyverse")
library("tidyverse")

install.packages("dplyr")
library("dplyr")

install.packages("tidyr")
library("tidyr")

install.packages("lubridate")
library(lubridate)

install.packages("ggplot2")
library(ggplot2)

install.packages("magrittr")
library(magrittr)
```

## Import the dataset
Choose relevant dataset and import them.
```{r}
setwd("C:/Users/HP/Desktop/GDA/Case study/Fitabase Data 4.12.16-5.12.16")
activity_log<- read.csv("dailyActivity_merged.csv")
sleep_log<- read.csv("sleepDay_merged.csv")
met_log<- read.csv("minuteMETsNarrow_merged.csv")
hourly_calories<- read.csv("hourlyCalories_merged.csv")
hourly_intensity<- read.csv("hourlyIntensities_merged.csv")
hourly_steps<-read.csv("hourlySteps_merged.csv")
```
##What's in the data
Get a glimpse of the data to see the number of columns, name of the columns, their data type, number of rows,etc. Check the data for NA or duplicate values and remove any if found
```{r}
# Glimpse of the data.

glimpse(activity_log)
glimpse(sleep_log)
glimpse(met_log)
glimpse(hourly_calories)
glimpse(hourly_intensity)
glimpse(hourly_steps)

# Check for NA values.

sum(is.na(activity_log))
sum(is.na(hourly_calories))
sum(is.na(hourly_intensity))
sum(is.na(hourly_steps))
sum(is.na(met_log))
sum(is.na(sleep_log))

# Check for duplicate values. 

sum(duplicated(activity_log))
sum(duplicated(hourly_calories))
sum(duplicated(hourly_intensity))
sum(duplicated(hourly_steps))
sum(duplicated(met_log))
sum(duplicated(sleep_log))

# Remove the duplicated values.

sleep_log<-sleep_log[!duplicated(sleep_log),]

# Check the data type of each column in the data Sets.

str(activity_log)
str(sleep_log)
str(met_log)
str(hourly_calories)
str(hourly_intensity)
str(hourly_steps)

# All the date fields are in chr.Convert it into datetime.

mdy<-mdy(activity_log$ActivityDate)
activity_log$ActivityDate<-mdy

mdy_hms<-mdy_hms(sleep_log$SleepDay)
sleep_log$SleepDay<-mdy_hms

met_log$ActivityMinute<-parse_date_time(met_log$ActivityMinute,'%m/%d/%Y %I:%M:%S %p')
hourly_intensity$ActivityHour<-parse_date_time(hourly_intensity$ActivityHour,'%m/%d/%Y %I:%M:%S %p')
hourly_calories$ActivityHour<-parse_date_time(hourly_calories$ActivityHour,'%m/%d/%Y %I:%M:%S %p')
hourly_steps$ActivityHour<-parse_date_time(hourly_steps$ActivityHour,'%m/%d/%Y %I:%M:%S %p')


# Converting date-time column into two separate columns. 

met_log<-tidyr::separate(met_log,ActivityMinute,c("Date","Time"),sep=" ")
```
## Addition of new columns
Added the day of the week column to get better understanding of the data.Also combined few columns in MET data.New columns to get the duration for which the device was worn, so that it can be used in the analysis.
```
{r}
# Calculate the total met value on daily basis.

met_log_daily<-met_log %>%
    group_by(Id,Date)%>%
  summarise(Total_Met=sum(METs ,na.rm = TRUE))
  
# To work with data better we add one new column 'Day'.

activity_log <- activity_log %>% mutate( Day = weekdays(as.Date(ActivityDate, "%m/%d/%Y")))

# Columns to identify the time(in minutes and hours) for which the device was worn.

activity_log$Worn_min<-activity_log$VeryActiveMinutes+activity_log$FairlyActiveMinutes+activity_log$LightlyActiveMinutes+activity_log$SedentaryMinutes
activity_log$Worn_hour<-activity_log$Worn_min/60
activity_log$worn_all_day<-activity_log$Worn_hour==24

worn_day<-count(activity_log,worn_all_day)%>%
  rename(Count=n)

```
## Analyse the data
Developed graphs to get an insight of the trends or patterns in the data.Summary of the data gives the average and also the min and max points.
```{r}
# Summary of the data to see their minimum, maximum and average values.

activity_log %>%
  select(TotalSteps,TotalDistance,Calories,Worn_hour,VeryActiveDistance,VeryActiveMinutes,LightActiveDistance,LightlyActiveMinutes,ModeratelyActiveDistance,
         SedentaryActiveDistance,SedentaryMinutes,FairlyActiveMinutes) %>%
  summary()
sleep_log %>%
  select(TotalMinutesAsleep,TotalTimeInBed)%>%
  summary()
met_log_daily%>%
  summary(Total_Met)

# Sedentary minutes vs weekday.

ggplot(activity_log,aes(x=Day,y=SedentaryMinutes,fill=Day))+
  geom_bar(stat = "identity",fill="#B87333")+
  labs(title="Sedantary minutes for day of the week",x="day of the week",y="Sedantary minutes",caption="940 observations")

# Calorie vs Step taken

ggplot(activity_log,aes(y=Calories,x=TotalSteps))+geom_point()+
  geom_smooth()+
  labs(title="Calories burnt for the steps taken",x="Steps taken on a day",y="Calories burnt",caption="940 Observations")

# Calorie vs Distance

ggplot(activity_log,aes(x=TotalDistance,y=Calories))+geom_point()+geom_smooth()+
  labs(title="Calories burnt Vs the Distance covered",x="Distance covered in a day",y="Calories burnt",caption="940 Observations")

# Minutes asleep vs Total time in bed

ggplot(sleep_log,aes(x=TotalMinutesAsleep,y=TotalTimeInBed))+geom_point()+geom_smooth()+
  labs(title = "Time spent sleeping Vs Time spent on bed",x="Minutes asleep",y="Time spent on bed",caption="410 Observations")

# Distance vs Day of the week

activity_log$Day<-factor(activity_log$Day,levels =c("Sunday","Monday","Tuesday","Wednesday","Thursday","Friday","Saturday") )
ggplot(activity_log,aes(x=Day,y=TotalDistance,fill=Day))+geom_bar(stat = "identity",fill="#00FFFF")+
  labs(title="Distance overed in a day",x="Day of the week",y="Distance covered",caption="940 Observations")


# How many people wore the device for the whole day.?

ggplot(worn_day)+
  geom_col(mapping=aes(x=worn_all_day ,y=Count, fill=worn_all_day))+
  geom_text(mapping=aes(x=worn_all_day ,y=Count , label=Count),position = position_stack(vjust = 0.9))+
  labs(title = "Was the device worn all day.?",caption = "940 observations")+
  xlab("Trackers worn all day")+
  ylab("Count")

# Hours for which the device was worn on each day of the week.?

hours_per_day <- activity_log[c("Id","ActivityDate","Day","Worn_hour")]
hours_per_day$Day<-factor(hours_per_day$Day, levels = c("Sunday","Monday","Tuesday","Wednesday","Thursday","Friday","Saturday"))

ggplot(hours_per_day,aes(x=Day,y=Worn_hour,group=1))+
 geom_point(alpha=1/10)+
  geom_smooth(color="#00008B")+
  labs(title="Hours worn per day of the week",x="Day of the week",y="Hours",caption = "940 Observations")
  
# How many people wore on all days of the week.?

wore_all_day<-activity_log[c("Day","worn_all_day")] %>%
  count(Day,worn_all_day)
wore_all_day$Day<-factor(wore_all_day$Day,levels =c("Sunday","Monday","Tuesday","Wednesday","Thursday","Friday","Saturday") )

ggplot(wore_all_day,aes(fill=worn_all_day,x=Day,y=n,))+
  geom_bar(position = "stack",stat="identity")+
  geom_text(mapping = aes(x=Day,y=n,label=n),position=position_stack(vjust = 0.5))+
  labs(title="Number of people who wore the device for the whole day of the week",
       x="Day of the week",
       y="Count of the participants",
       caption="940 Observations")

```


